---
title: "Analysis of walking Steps taken by an individual during 2 months"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/home/amol/Projects/coursera/R/RepData_PeerAssessment1/")
```
## Loading and preprocessing the data

```{r}
data<-read.csv("activity.csv")
```
#We will look at data, column data type and stats about steps

```{r data, echo=TRUE}
head(data)
tail(data)
summary(data)
str(data)


```




##mean total number of steps tak

```{r data, echo=TRUE}
steps_per_day<-tapply(data$steps,data$date,function(x) sum(x,na.rm=TRUE))
mean(steps_per_day,na.rm=TRUE)
median(steps_per_day,na.rm=TRUE)
hist(steps_per_day)
```
-Mean of total steps per day = 9354
-Median of total steps per day = 10,395

##Daily activity pattern


```{r echo=TRUE}
steps_per_interval<-tapply(data$steps,data$interval,function(x) sum(x,na.rm=TRUE))
mean(steps_per_interval,na.rm=TRUE)
plot(names(steps_per_interval),steps_per_interval,type="l",xlab="interval",ylab="steps")
which(steps_per_interval==max(steps_per_interval))
```

- Heighest number of steps are observed in interval 835
##Missing Values

#Filling missing Values
```{r echo=TRUE}
data_filled<-data
mean_steps<-tapply(data$steps,data$interval,function(x) mean(x,na.rm = TRUE))
for (i in 1:length(data)){
    if (is.na(data[i,1])){
        data_filled[i,1]<-as.numeric(mean_steps[i])
    }
}
steps_per_day_filled <- tapply(data_filled$steps,data_filled$interval,function(x) sum(x,na.rm = TRUE))
hist(steps_per_day_filled,,xlab="steps_per_day")
mean(steps_per_day_filled,na.rm=TRUE)
median(steps_per_day_filled,na.rm=TRUE)
```
-Mean of total steps per day = 1981
-Median of total steps per day = 1808

The mean values have gone down by almost 5 times. 

##Are there differences in activity patterns between weekdays and weekends?

```{r}
data_filled$date<-as.Date(data_filled$date)
data_filled$day<-sapply(data_filled$date,weekdays)
f<-function(x){ if (x=='Sunday'|x=="Saturday")
    return ("Weekend") 
    else return ("Weekday")}
data_filled$day<-sapply(data_filled$day,f)


```


```{r}
par(c(2,1))
data_day<- data_filled%>%select(day,interval,steps)%>%group_by(day,interval)%>%summarize(mean=mean(steps,na.rm=TRUE))

data_day_weekday<-data_day%>%filter(day=='Weekday')
data_day_weekend<-data_day%>%filter(day=='Weekend')
head(data_day_weekend)
plot(data_day_weekday$interval,data_day_weekday$mean,xlab="interval",ylab="Avg. steps",main="Activity over Weekdays",type="l")
plot(data_day_weekend$interval,data_day_weekend$mean,xlab="interval",ylab="Avg. steps",main="Activity over Weekend",type="l")
```